Liveness-Based Garbage Collection for Lazy Languages

### 2015-07-03 paper/paper.tex
new abstract

% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.
\cmt{{
\numberofauthors{3} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors
section.
%
\author{
% 1st. author
\alignauthor K. Prasanna Kumar \\
       \affaddr{IIT Bombay,}\\
       \affaddr{Mumbai 400076, India}
       \email{prasanna@cse.iitb.ac.in}
% 2nd. author
\alignauthor Amey Karkare\\
       \affaddr{IIT Kanpur}\\
       \affaddr{Kanpur 208016, India}
       \email{karkare@cse.iitk.ac.in}
% 3rd. author
\alignauthor Amitabha Sanyal \\
       \affaddr{IIT Bombay}\\
       \affaddr{Mumbai 400076, India}\\
       \email{as@cse.iitb.ac.in}
}
}}
\authorinfo{Double Blind Review}{}{}
\maketitle

\begin{abstract}
We consider  the problem of reducing  the memory required  to run lazy
first-order functional programs.  To  do this, we analyze programs for
liveness of heap  allocated data, i.e. data that  can possibly be used
in the future.  The accompanying garbage collector is  modified to use
the result of the analysis to preserve only {\em live} data---a subset
of {\em reachable data}---during  garbage collection.  This results in
an  increase in  the garbage  reclaimed and  consequently  reduces the
memory footprints of programs.   While this technique has already been
shown to yield benefits  for eager first-order languages, the presence
of  closures   at  run-time  poses  additional   challenges  for  lazy
languages.   Closures require  changes  both in  the earlier  liveness
analysis for heap allocated memory  and also in the garbage collection
scheme required to reclaim memory that is not live.

We  formulate  a context-sensitive  liveness  analysis  and using  the
results annotate each potential allocation point in the program with a
set     of    deterministic    finite-state     automata(DFA).     The
garbage-collector  inspects  the DFA  to  curtail reachability  during
marking.  As a result, fewer  objects are marked (with a possibly more
expensive  marker)  and  preserved  by the  copy  phase.   Experiments
confirm  that  liveness-based  collection  results in  a  increase  in
garbage  reclaimed  and  a   consequent  decrease  in  the  number  of
collections. On  our benchmark programs, we obtain  a decrease between
x\% and y\% in the memory size required to run programs, when compared
to reachability-based garbage collection.
\end{abstract}

>>>>

% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.
\cmt{{
\numberofauthors{3} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors
section.
%
\author{
% 1st. author
\alignauthor K. Prasanna Kumar \\
       \affaddr{IIT Bombay,}\\
       \affaddr{Mumbai 400076, India}
       \email{prasannak@cse.iitb.ac.in}
% 2nd. author
\alignauthor Amey Karkare\\
       \affaddr{IIT Kanpur}\\
       \affaddr{Kanpur 208016, India}
       \email{karkare@cse.iitk.ac.in}
% 3rd. author
\alignauthor Amitabha Sanyal \\
       \affaddr{IIT Bombay}\\
       \affaddr{Mumbai 400076, India}\\
       \email{as@cse.iitb.ac.in}
}
}}
\authorinfo{Double Blind Review}{}{}
\maketitle

\begin{abstract}
We consider  the problem of reducing  the memory required to  run lazy
first-order functional programs.  To do  this, we analyze programs for
{\em liveness} of heap allocated data, i.e.  data that can possibly be
used in the future.  The accompanying garbage collector is modified to
use the  result of the analysis  to preserve only {\em  live} data---a
subset  of {\em  reachable  data}---during  garbage collection.   This
results  in an  increase  in the  garbage  reclaimed and  consequently
reduces the memory  footprints of programs.  While  this technique has
already been shown to yield  benefits for eager first-order languages,
the presence of  closures at run-time poses  additional challenges for
lazy languages.  Closures require changes both in the earlier liveness
analysis for heap allocated memory  and also in the garbage collection
scheme required to reclaim memory that is not live.

Our  analysis  is a  context-sensitive  analysis  that annotates  each
potential  garbage collection  point  in  the program  with  a set  of
deterministic finite-state  automata (DFA) describing the  liveness at
that  point.   We  also  present a  practical  implementation  of  our
analysis in  a garbage collector.   The collector, when  invoked, uses
the DFA set to preserve live  objects.  As a result, fewer objects are
marked and preserved when compared  to a reachability based marking.
Our experiments  confirm that liveness-based collection  increases the
garbage reclaimed,  consequently decreasing the number  of collections
on many  practical programs.  On  our benchmark programs, we  obtain a
decrease of  up to  y\% in  the number  of collections,  and up  to x\%
decrease in overall garbage collection time.
\end{abstract}
